Wide Post-Mass Transfer White-Dwarf Binaries

### 2024-07-08 summaries/paper-draft-plain.tex
modified high mass unstable part of summary and reran relevant code

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[top=0.60in, bottom=0.80in, left=0.60in, right=0.60in]{geometry}
\usepackage{graphicx}
\usepackage{siunitx}
\usepackage{listings}

\newcommand{\Msun}{\mathrm{M_{\odot}}}
\newcommand{\Rsun}{\mathrm{R_{\odot}}}
\newcommand{\MWD}{M_{\mathrm{WD}}}
\newcommand{\Mstar}{M_{\star}}
\newcommand{\alphace}{\alpha_{\mathrm{CE}}}
\newcommand{\alphath}{\alpha_{\mathrm{th}}}
\newcommand{\alpharec}{\alpha_{\mathrm{rec}}}
\newcommand{\Ebind}{E_{\mathrm{bind}}}
\newcommand{\Egrav}{E_{\mathrm{grav}}}
\newcommand{\Eint}{E_{\mathrm{int}}}
\newcommand{\Eth}{E_{\mathrm{th}}}
\newcommand{\Erec}{E_{\mathrm{rec}}}
\newcommand{\Porb}{P_{\mathrm{orb}}}
\newcommand{\au}{\, \mathrm{au}}

@@

In this paper, we are going to explore the formation of these wide post-mass transfer WD+MS binaries with COSMIC model. We perform COSMIC simulation on both individual binary star systems and sampled population to explore the formation of WD+MS binaries under different conditions. In section \ref{sec:ce}, we are going to investigate formation of individual wide WD+MS binaries through CE for a variety of variables, including initial mass, initial separation, CE efficiency, and the energy budget of the CE. In section \ref{sec:stable}, we are going to explore the formation of individual wide WD+MS binaries through stable mass transfer. In section \ref{sec:population}, we will research on how different COSMIC model affect simulation results of a sample population of binary star systems, and how the formation of post WD+MS binaries depends on various parameters.

\section{Common Envelope} \label{sec:ce}
In \cite{yamaguchi_hi, yamaguchi_lo}, the author discussed the formation of WD+MS binaries through CE for both high-mass and low-mass systems, and it was found that wide WD+MS systems can be formed for certain energy budgets and certain initial separation in the MESA model. We hope to explore similar models in COSMIC.

@@

In COSMIC, the CE is modeled using a structural parameter $\lambda$, where
\begin{equation}
  \Ebind = \frac{G M_i M_{\mathrm{env}}}{\lambda R_i}.   
  \label{ebind}
\end{equation}
This $\lambda$ is represented in COSMIC as the \verb|lambdaf| flag in \verb|BSEDict|. To compare MESA model with COSMIC, we intend to compare how the energy budget utilized by \cite{yamaguchi_hi, yamaguchi_lo} matches with the CE model in COSMIC. We approach this by calculating the effective $\lambda$ value for each of the energy budget used in \cite{yamaguchi_hi, yamaguchi_lo}.

@@

In \cite{yamaguchi_hi}, the lower limit of WD mass $M_{\mathrm{MD, min}}$ in the observational sample is calculated to be $1.244\Msun \sim 1.418\Msun$. The corresponding mass of MS progenitor is expected to be in the range of $6 \sim 9 \Msun$, and the companion mass has median around $1\Msun$. MESA results for $7\Msun + 1\Msun$ systems is documented in \cite{yamaguchi_hi}. Hence, we also consider a $7\Msun + 1\Msun$ evolution model in COSMIC and compare the results with those in \cite{yamaguchi_hi}.

We run COSMIC binary evolution simulation of a $7\Msun + 1\Msun$ system and investigate how the final separation $a_f$ depends on the initial separation $a_i$ and the CE parameter $\lambda$. We vary the initial separation in range $2 \sim 8 \au$ with $400$ steps. We vary the CE flag \verb|lambdaf| in range $0 \sim -100$, which corresponds to $\lambda = 0 \sim 100$, with $2000$ steps. Also, we run these parameters with four different CE efficiencies — $\alphace = 1$, $\alphace = 0.9$, $\alphace = 0.6$, and $\alphace = 0.3$, the same as documented in \cite{yamaguchi_hi}. In total, for $4$ different CE efficiency $\alphace$, we run a grid of $80000$ binary star systems with different initial separation $a_i$ and CE parameter $\lambda$.

After simulation is finished, we select the time-step when the system finished CE for the first time (\verb|evol_type = 8|). At this moment, if the system is a WD+MS system, we record their separation as the final separation $a_f$. Otherwise, we set $a_f = 0$ and consider this initial condition unable to produce a WD+MS system.

We create four heat maps for the four different CE efficiencies. In Figure \ref{res_hi}, we plot the final separation $a_f$ of the WD+MS systems against initial separation $a_i$ and the CE flag \verb|lambdaf|$= - \lambda$.

\begin{figure}
  \centering
  \includegraphics[width=0.8\linewidth]{fig/7+1results.png}
  \caption{ Heat map of final separation $a_f$ against initial separation $a_i$ and CE flag lambdaf$=-\lambda$. Different panels represent four different CE efficiencies $\alphace = 1$, $0.9$, $0.6$, and $0.3$. The white and yellow contour corresponds to $a_f = 0.01 \au$ and $a_f = 0.15 \au$ respectively. WD+MS binaries is not possible for separation smaller than $0.01 \au$, and $0.15 \au$ is the minimum separation of wide WD+MS binaries in observational results. Outside the white contour no desired system forms.}
  \label{res_hi}
\end{figure}

@@

There are several interesting points of the results. First of all, as $\alphace$ decreases, the final separation decreases in general. This can be explained by Equation \ref{ebind-sep}. Qualitatively, note that more energy from the orbit is needed to eject the envelope for low $\alphace$, so the resulting $a_f$ is expected to be smaller.

It is worth noting that most of these systems experience two mass transfer phases in COSMIC before reaching WD+MS. That is, during the evolution process, two CE processes take place, between which there is a period of time. By checking the \verb|bpp| array, which documents the evolution processes of the binary star systems, we found that (***TO-DO***) fraction of all our simulated binaries experience two CE processes before reaching WD+MS. When the primary loses too much mass during the first CE, the CE process ends. After the primary fills its Roche lobe and has a large enough envelope again, the CE process resumes and causes two CE processes in total.

From Figure \ref{res_hi}, we also notice that the final separation $a_f$ jumps obviously at about $2.9 \au$, and $7.1 \au$. After investigating the \verb|bpp| array of these simulations, we find out the jump at about $2.9 \au$ and $7.1 \au$ is likely due to the difference of \verb|kstar_1| at the start of RLOF. For $a_i < 2.9 \au$, RLOF starts when the primary is on the giant branch (GB) with \verb|kstar_1 3|. For $a_i = 2.9 \au$, RLOF starts when the primary is experiencing core Helium burning (CHeB) with \verb|kstar_1 = 4|. For $2.9 \au < a_i < 7.1 \au$, RLOF starts when the primary is on E-AGB with \verb|kstar_1 = 5|. For $a_i > 7.1 \au$, RLOF starts when the primary is on TP-AGB with \verb|kstar_1 = 6|. For different star type, the structure and mass of the envelope is different, leading to different binding energy $\Ebind$ and thus different final separation $a_f$.

\begin{figure}
  \centering
  \includegraphics[width = 0.5\linewidth]{fig/jump-zoom.png}
  \caption{Dependence of final separation $a_f$ on initial separation $a_i$ for $\alphace = 0.3$. Simulation results of $\lambda_f = 60$, $70$, and $80$ are shown. The red circles emphasizes the jump of final separation as initial separation increases.}
  \label{jump-zoom}
\end{figure}

Inspecting more closely, we find that the final separation $a_f$ jumps at about $5.2 \sim 5.4 \au$. This is especially obvious with $\alphace = 0.3$. To illustrate, we zoom in the simulation results of $\alphace = 0.3$, and present the change of final separation $a_f$ as initial separation $a_i$ increases in Figure \ref{jump-zoom}. From here, we can clearly see two jumps — one between $5.1\au$ and $5.2\au$, the other at $5.34\au$. To explain this, we again go to the \verb|bpp| array of these systems. We notice that the first jump is probably because the evolution process changes from \textbf{3-7-8-4-3-7-8-4} to \textbf{3-7-8-7-8-4}. This means that for relatively larger initial separation $a_i$, the primary fills its Roche lobe all the time, leading to more mass loss during a shorter timescale and thus larger final separation. Also, the second jump is probably because \verb|kstar_1| changes from \verb|8| (Naked Helium Star Hertzsprung Gap) to \verb|9| (Naked Helium Star Giant Branch) at the start of the second common envelope. This changes the binding energy of the common envelope and thus changes the final separation.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.9\linewidth]{fig/7+1fit.png}
  \caption{Effective $\lambda$ fitting results of a $7\Msun + 1\Msun$ system for common envelope structure in MESA. Left panel: MESA results in \cite{yamaguchi_hi} of how final separation $a_f$ depends on initial separation $a_i$. Central panel: COSMIC results of how final separation $a_f$ depends on initial separation $a_i$, produced by the effective $\lambda$ calculated. Right panel: effective $\lambda$ in COSMIC that corresponds to the MESA results for each initial separation $a_i$. The red dashed line in the left and middle panel represents $0.15 \au$, which is the minimal separation for observed wide post-CE WD+MS binaries.}
  \label{fit_hi}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=0.9\linewidth]{fig/7+1cmp.png}
  \caption{Same as Figure \ref{fit_hi}, but with default $\lambda$ in COSMIC model included in the right panel.}
  \label{fit_cmp_hi}
\end{figure}

To further compare with the default $\lambda$ value in COSMIC, we include in the right panel an extra line which represents the default $\lambda$ value in COSMIC. This default $\lambda$ in the COSMIC model is calculated following Appendix A of \cite{claeys2014theoretical}. For all of our case, we have $M_{\mathrm{env}} > 1$. The results is shown in Figure \ref{fit_cmp_hi}. Notice that the default $\lambda$ value is at order $\sim 1$, far smaller than the effective $\lambda$ needed to recreate results in \cite{yamaguchi_hi}.

\begin{figure}
  \centering
  \includegraphics[width=0.9\linewidth]{fig/7+1ebcmp.png}
  \caption{Same as Figure \ref{fit_cmp_hi}, but with a different energy budget — $\Ebind = \Egrav + \alphath \Eth + \alpharec \Erec$.}
  \label{fit_cmp_eb_hi}
\end{figure}

In Figure \ref{fit_cmp_eb_hi}, we present the same results for another energy budget of the CE process — $\Ebind = \Egrav + \alphath \Eth + \alpharec \Erec$.

@@

\section{Stable Mass Transfer} \label{sec:stable}

\section{Population} \label{sec:population}

\bibliography{reference.bib}
\bibliographystyle{apalike}

>>>>

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[top=0.60in, bottom=0.80in, left=0.60in, right=0.60in]{geometry}
\usepackage{graphicx}
\usepackage{siunitx}
\usepackage{listings}
\usepackage[labelfont=bf]{caption}

\newcommand{\Msun}{\,\mathrm{M_{\odot}}}
\newcommand{\Rsun}{\,\mathrm{R_{\odot}}}
\newcommand{\MWD}{M_{\mathrm{WD}}}
\newcommand{\Mstar}{M_{\star}}
\newcommand{\alphace}{\alpha_{\mathrm{CE}}}
\newcommand{\alphath}{\alpha_{\mathrm{th}}}
\newcommand{\alpharec}{\alpha_{\mathrm{rec}}}
\newcommand{\Ebind}{E_{\mathrm{bind}}}
\newcommand{\Egrav}{E_{\mathrm{grav}}}
\newcommand{\Eint}{E_{\mathrm{int}}}
\newcommand{\Eth}{E_{\mathrm{th}}}
\newcommand{\Erec}{E_{\mathrm{rec}}}
\newcommand{\Porb}{P_{\mathrm{orb}}}
\newcommand{\au}{\, \mathrm{au}}

@@

In this paper, we are going to explore the formation possibility of these wide post-mass transfer WD+MS binaries with COSMIC model. We perform COSMIC simulation on both individual binary star systems and sampled population to explore the formation of WD+MS binaries under different conditions. In section \ref{sec:ce}, we are going to investigate formation of individual wide WD+MS binaries through CE for a variety of variables, including initial mass, initial separation, CE efficiency, and the energy budget of the CE. In section \ref{sec:stable}, we are going to explore the formation of individual wide WD+MS binaries through stable mass transfer. In section \ref{sec:population}, we will research on how different COSMIC model affect simulation results of a sample population of binary star systems, and how the formation of post WD+MS binaries depends on various parameters.

\section{Formation Through Common Envelope} \label{sec:ce}
In \cite{yamaguchi_hi, yamaguchi_lo}, the author discussed the formation of WD+MS binaries through CE for both high-mass and low-mass systems, and it was found that wide WD+MS systems can be formed for certain energy budgets and certain initial separation in the MESA model. We hope to explore similar models in COSMIC.

@@

In COSMIC, the CE is modeled using a structural parameter $\lambda$, where
\begin{equation}
  \Ebind = \frac{G M_i M_{\mathrm{env}}}{\lambda R_i}.   
  \label{ebind}
\end{equation}
This $\lambda$ is represented in COSMIC as the \verb|lambdaf| flag in \verb|BSEDict|. To compare MESA model with COSMIC, we intend to compare how the energy budget utilized by \cite{yamaguchi_hi, yamaguchi_lo} matches with the CE model in COSMIC. We approach this by calculating the effective $\lambda$ value for each of the common envelope energy budget used in \cite{yamaguchi_hi, yamaguchi_lo}.

@@

According to \cite{yamaguchi_hi}, the lower limit of WD mass $M_{\mathrm{MD, min}}$ in the observational sample is calculated to be $1.244\Msun \sim 1.418\Msun$. The corresponding mass of MS progenitor is expected to be in the range of $6 \sim 9 \Msun$, and the companion mass has median around $1\Msun$. MESA results for $7\Msun + 1\Msun$ systems is documented in \cite{yamaguchi_hi}. Hence, we also consider a $7\Msun + 1\Msun$ evolution model in COSMIC and compare the results with those in \cite{yamaguchi_hi}.

We run COSMIC binary evolution simulation of a $7\Msun + 1\Msun$ system and investigate how the final separation $a_f$ depends on the initial separation $a_i$ and the CE parameter $\lambda$. The initial separation is a linear grid in range $2 \sim 8 \au$ with $400$ points. The CE flag \verb|lambdaf| is a linear grid in range $0 \sim -100$ with $2000$ steps, which corresponds to $0 \sim 100$ for CE parameter $\lambda$. Also, we run these parameters with four different CE efficiencies — $\alphace = 1$, $\alphace = 0.9$, $\alphace = 0.6$, and $\alphace = 0.3$, the same as recorded in \cite{yamaguchi_hi}. In total, for each of the $4$ different CE efficiency $\alphace$, we run a grid of $80000$ binary star systems with different initial separation $a_i$ and CE parameter $\lambda$.

After simulation is finished, we select the time-step when the system finished CE for the first time (\verb|evol_type=8|). At this moment, if the system is a WD+MS system, we record their separation as the final separation $a_f$. Otherwise, we set $a_f = 0$ and consider this initial condition unable to produce a WD+MS system. We create four heat maps for the four different CE efficiencies. In Figure \ref{res_hi}, we plot the final separation $a_f$ of the WD+MS systems against initial separation $a_i$ and the CE flag $- \lambda$.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{fig/7+1results.png}
  \caption{ Heat map of final separation $a_f$ against initial separation $a_i$ and CE flag $-\lambda$. Different panels represent four different CE efficiencies $\alphace = 1$, $0.9$, $0.6$, and $0.3$. The white and yellow contour corresponds to $a_f = 0.01 \au$ and $a_f = 0.15 \au$ respectively. WD+MS binaries is not possible for separation smaller than $0.01 \au$, and $0.15 \au$ is the minimum separation of wide WD+MS binaries in observational results. Outside the white contour no desired system forms.}
  \label{res_hi}
\end{figure}

@@

There are several interesting points of the results. First of all, comparing the four panels, we found that as $\alphace$ decreases, the final separation decreases in general. This can be explained by Equation \ref{ebind-sep}. Qualitatively, note that more energy from the orbit is needed to eject the envelope for low $\alphace$, so the resulting $a_f$ is expected to be smaller.

It is worth noting that most of these systems experience two mass transfer phases in COSMIC before reaching WD+MS. That is, during the evolution process, two CE processes take place, between which there is a period of time. A typical change of \verb|evol_type| through the evolution is \textbf{\dots-3-7-8-4-\dots-3-7-8-4-\dots}, where WD+MS is reached during the second mass transfer process. Another possible change of \verb|evol_type| through the volution is \textbf{\dots-3-7-8-7-8-4-\dots}, which will be mentioned in the next paragraph. Statistically, we found that \textbf{(***TO-DO***)} fraction of all our $320000$ simulated binaries experience two CE processes before reaching WD+MS. To explain this qualitatively, we notice that when the primary loses too much mass during the first CE, the CE process ends, as the radius of the primary becomes smaller than the Roche Lobe. After the primary fills its Roche lobe and has a large enough envelope again, the CE process resumes and causes two CE processes in total.

From Figure \ref{res_hi}, we also notice that final separation $a_f$ jumps obviously at about $2.9 \au$, and $7.1 \au$. After investigating the \verb|bpp| array of these simulations, we find out the jump at about $2.9 \au$ and $7.1 \au$ is likely due to the difference of \verb|kstar_1| at the start of RLOF. For $a_i < 2.9 \au$, RLOF starts when \verb|kstar_1=3|. For $a_i = 2.9 \au$, RLOF starts when \verb|kstar_1=4|. For $2.9 \au < a_i < 7.1 \au$, RLOF starts when \verb|kstar_1=5|. For $a_i > 7.1 \au$, RLOF starts when \verb|kstar_1=6|. For different star type, the structure and mass of the envelope is different, leading to different binding energy $\Ebind$ and thus different final separation $a_f$.

\begin{figure}
  \centering
  \includegraphics[width = 0.6\linewidth]{fig/jump-zoom.png}
  \caption{Dependence of final separation $a_f$ on initial separation $a_i$ for $\alphace = 0.3$. Simulation results of $\lambda_f = 60$, $70$, and $80$ are shown. The red circles emphasizes the jump of final separation as initial separation increases.}
  \label{jump-zoom}
\end{figure}

Inspecting more closely, we found that the final separation $a_f$ jumps at about $5.2 \sim 5.4 \au$. This is especially obvious with $\alphace = 0.3$. To illustrate, we zoom in the simulation results of $\alphace = 0.3$, and present the change of final separation $a_f$ as initial separation $a_i$ increases in Figure \ref{jump-zoom}. From here, we can clearly see two jumps — one between $5.1\au$ and $5.2\au$, the other at $5.34\au$. To explain this, we again go to the \verb|bpp| array of these systems. The first jump is because the evolution process changes from \textbf{3-7-8-4-3-7-8-4} to \textbf{3-7-8-7-8-4}. This means that for relatively larger initial separation $a_i$, the primary fills its Roche lobe all the time, leading to more mass loss during a shorter timescale and thus larger final separation. For the second jump, it is because \verb|kstar_1| changes from \verb|8| to \verb|9| at the start of the second common envelope. This changes the binding energy of the common envelope and thus changes the final separation.

@@

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{fig/7+1fit.png}
  \caption{Effective $\lambda$ fitting results of a $7\Msun + 1\Msun$ system for common envelope structure in MESA. Different lines represent different $\alphace$ and different common envelope energy budget, as shown in the legend. Left panel: MESA results in \cite{yamaguchi_hi} of how final separation $a_f$ depends on initial separation $a_i$. Central panel: COSMIC results of how final separation $a_f$ depends on initial separation $a_i$, produced by the effective $\lambda$ calculated. Right panel: effective $\lambda$ in COSMIC that corresponds to the MESA results for each initial separation $a_i$. The red dashed line in the left and middle panel represents $0.15 \au$, which is the minimal separation for observed wide post-CE WD+MS binaries.}
  \label{fit_hi}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{fig/7+1cmp.png}
  \caption{Same as Figure \ref{fit_hi}, but with default $\lambda$ in COSMIC model included in the right panel.}
  \label{fit_cmp_hi}
\end{figure}

To further compare with the default $\lambda$ value in COSMIC, we include in the right panel an extra line which represents the default $\lambda$ value in COSMIC. This default $\lambda$ in the COSMIC model is calculated following Appendix A of \cite{claeys2014theoretical}. For all of our case, we have $M_{\mathrm{env}} > 1$. The results is shown in Figure \ref{fit_cmp_hi}. Notice that the default $\lambda$ value is at order $\sim 1$, far smaller than the effective $\lambda$ needed to recreate results in \cite{yamaguchi_hi}. Hence, we conclude that in COSMIC, it is not practical to form wide high mass WD+MS system through common envelope processes, since the required $\lambda$ is too large.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{fig/7+1ebcmp.png}
  \caption{Same as Figure \ref{fit_cmp_hi}, but with a different common envelope energy budget. For results shown in these figures, $\Ebind = \Egrav + \alphath \Eth + \alpharec \Erec$, where the coefficients $\alphath$ and $\alpharec$ for different line is shown in the legend.}
  \label{fit_cmp_eb_hi}
\end{figure}

In Figure \ref{fit_cmp_eb_hi}, we present the same results for another energy budget of the CE process. Rather than including both thermal and recombination energy into the binding energy, we use two coefficients $\alphath$ and $\alpharec$ to consider their contribution to $\Ebind$ separately. That is,
\[
  \Ebind = \Egrav + \alphath \Eth + \alpharec \Erec.
\]
We notice that the effective $\lambda$ now gets with in the order of $\sim 10$. However, none of these initial separation and $\lambda$ results in wide WD+MS system with final separation $a_f > 0.15 \au$. Therefore, we conclude that with the new energy budget, it is not practical to form wide WD+MS binaries either.

@@

\section{Formation Through Stable Mass Transfer} \label{sec:stable}

\section{Population Synthesis Results} \label{sec:population}

\bibliography{reference.bib}
\bibliographystyle{alpha}
